### 2024-01-23 outputs/paper/paper.qmd
General format

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-01-23 outputs/paper/paper.qmd
Updated

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
+title: "Analyzing Trends in Crime Against Persons: Insights from the Toronto Police Annual Statistical Report"
+author: "Mohammed Yusuf Shaikh"
 thanks: "Code and data are available at: LINK."
 date: today
-date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+date-format: 22 January 2024
+abstract: "Public Safety and security are a paramount concern in the 20th century, the analysis of statistics in Victimology helps to visualize and derive the conclusions of urban crime dynamics.
+This analytical report delves into the trends of victims of a certain crime in Toronto from the year 2014 to 2022.
+Through a series of data visualization approach to examine the relationship between crime incidents and demographic factors.
+The data reveals significant fluctuations in crime rates in correspond to the socio-economic shifts, including a notable dip in crimes reported during pandemic between 2019-2020.
+The visualization of the certain trends provides a compelling narrative on the dynamics of victims of crime in Toronto, advocating for all the policymakers to call for a reevaluation of current approaches to public safety and crime prevention."
 format: pdf
@@
 
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
 ```
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
 ```
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-01-23 scripts/02-data_cleaning.R
Commit for scripts

@@ -37,7 +37,7 @@ head(cleaned_victim_data)
 
 
 # Save data in inputs/data
-write_csv(cleaned_victim_data, file= "inputs/data/cleaned_data.csv")
+write_csv(cleaned_victim_data, file="outputs/data/cleaned_lead_data.csv")
 
 
 

### 2024-01-23 scripts/03-test_data.R

@@ -28,7 +28,11 @@ cleaned_victim_data$report_year |> min() == 2014
 # Test 3 check maximum year in data
 cleaned_victim_data$report_year |> max() == 2022
 
-# Testing class for victim_count data
+# Test 4 Tests that there are exactly 1110 id in the data
+num_rows <- nrow(cleaned_victim_data)
+num_rows == 1110
+
+# Test 5 Testing class for victim_count data
 cleaned_victim_data$victim_count |> class() == "numeric"
 
 

### 2024-01-23 outputs/paper/paper.qmd
Upload

@@
 date: today
-date-format: 22 January 2024
+date-format: long
 abstract: "Public Safety and security are a paramount concern in the 20th century, the analysis of statistics in Victimology helps to visualize and derive the conclusions of urban crime dynamics.
@@
 library(tidyverse)
-library(palmerpenguins)
+library(AER)
+library(opendatatoronto)
+library(dplyr)
+library(readr)
+library(janitor)
+library(ggplot2)
+
+# Loading Clean data & Read data
+cleaned_victim_data <- read_csv(cleaned_data, file="outputs/data/cleaned_lead_data.csv")
 ```
@@
 # Introduction
+The concern of Public safety in the Urban society is a necessary step towards safeguarding people of that community.
+The city of Toronto has various socio-economic challenges one of them being Crime against a person.
+A crime incident might leave a victim and our report aims to explore the patterns and trend of crimes against the person.
+The paper aims to understand how this trends have evolved among victims and analyzing there  demograpics from 2014 to 2022.
+To elaborate, the report context is not only quantifying crime victims but also interpret and explain whereby it motivates to a better approach towards urban safety and policies.
+The Criminal victimization is a frightening and unsettling experience for many Canadians (Cite).
 
-You can and should cross-reference sections and sub-sections.
+Victimology is a vast sub-category of criminology to understand various emotional factors and its causation, however based on our police annual report – victims of  crime showcase a statistical record based on which the report will share the knowledge to work on the gap to understand between victim demographics and crime.
+This report addresses this gap by visualizing crime statistics through a demographic lens, thereby unveiling patterns that may not be evident in numerical data alone
 
-The remainder of this paper is structured as follows. @sec-data....
+Through R programming language [@R] we analyze the data obtained from Open data Toronto [Open data Toronto].
+This dataset provides a detailed account of all recorded count of victim under the jurisdiction of the Toronto Police Services over an eight-period.
+Our research offers an examination of cleaned data points and produced visual formats that reveal the shapes the crime type as well as their victims across different genders and age cohorts.
+The analysis, supplemented with various analytical tools like Tidyverse [@tidyverse], Janitor [@janitor], and Knitr [@knitr], delves deep into the trends, revealing patterns that could have implications for policy formulation and implementation.
 
-# Data {#sec-data}
+The Introduction is followed by the Data section.
+Within the data section we talk about the data source used for this report, the data cleaning process applied and the result that is the visual analysis of trends and pattern involved.
+The Discussion section shares additional insights into the data findings, as well as discussing measurement errors and limitations of this analysis.
+Lastly, the document summarizes the analysis.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-```
-
-\newpage
+# Data
 

### 2024-01-23 scripts/02-data_cleaning.R

@@ -37,7 +37,7 @@ head(cleaned_victim_data)
 
 
 # Save data in inputs/data
-write_csv(cleaned_victim_data, file="outputs/data/cleaned_lead_data.csv")
+write_csv(cleaned_victim_data, file="outputs/data/cleaned_data.csv")
 
 
 

### 2024-01-24 outputs/paper/paper.qmd
Updated File

@@
 library(ggplot2)
+library(knitr)
 
-# Loading Clean data & Read data
-cleaned_victim_data <- read_csv(cleaned_data, file="outputs/data/cleaned_lead_data.csv")
+raw_data <- read_csv(here::here("inputs/data/raw_data.csv"))
+
+cleaned_dataset <-
+  read_csv(
+    here::here("outputs/data/cleaned_data.csv")
+  )
 ```
@@
 
-# Data
+# Data {#sec-2}
 
-# References
+## Data Source
+
+The data used in this paper was collected by the OpenDataToronto Library [@citeODT]. The specific dataset used in this research is the
+'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset] from the year 2014 to 2022.
+The dataset is categorized by crime subtype against a person, so basically four-types: Assault, Robbery, Sexual Violation, and Other offenses.
+The stratification of the dataset follows victim demographics such as age group, age cohort and gender.
+
+## Data Cleaning
+```{r}
+#| label: tbl-Cleaned-data-output
+#| tbl-cap: Sample of Cleaned Victim Dataset 
+#| echo: false
+head(cleaned_dataset) |>
+  kable(
+    col.names = c("ID", "YEAR", "CRIME", "SEX", "AGE_GROUP", "AGE_COHORT", "VICTIM_COUNT"),
+    booktabs = TRUE
+    )
+```
+
+Data used in this paper was downloaded, cleaned and analyzed with the programming language R [@citeR].
+Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr].
+The Table 1 showcases the cleaned dataset from the raw data extraced from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].
+
+# Results & Analysis
+
+## Total Victims of Crime over Year
+
+Refer to @tbl-Total-Victims-by-Year which shows the data on totals victims of crime over the period of 2014 to 2022 and corresponding to that a line graph is created, both showcase the annual trend in the number of reported victims.
+@tbl-Total-Victims-by-Year, titled "Total Victims From 2014-2022", is shown in the tabular format, allowing for quick reference to the absolute numbers of victims per year.
+
+```{r}
+#| label: tbl-Total-Victims-by-Year
+#| tbl-cap: Total Victims From 2014-2022
+#| echo: false
+summarized_data <- cleaned_dataset %>%
+  group_by(report_year) %>%
+  summarize(Total_Count = sum(victim_count))
+
+# Use kable from knitr to create a simple table
+
+summarized_data |>
+  kable(
+    col.names = c("YEAR", "TOTAL VICTIMS OF CRIME"),
+    booktabs = TRUE
+    )  
+```
+
+```{r}
+#| label: fig-ttl-victims
+#| fig-cap: Total Victims of Crime over Year (2014-2022)
+#| echo: false
+#| warning: false
+#| message: false
+
+# Creating the line graph
+ggplot(summarized_data, aes(x = report_year, y = Total_Count)) +
+  geom_line(color = "blue") +  # This creates the line graph
+  geom_point() +  # Optional: adds points at each year
+  labs(title = "Total Victims of Crime",
+       x = "Year",
+       y = "Total Victims of Crime Over Year") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
+
+```
+
+From @fig-ttl-victims - the line graph provides a visual representation of these figures, plotting the total victims of crime over the years.
+The trend shows a steady increase in Total victim counts from 2014 to 2019, suggesting an increasing rate of victimization during this period and increasing crime rate.
+However, there is a sharp decline in the year 2020 that invites questions about the potential causes.
+That was followed by an upward trajectory in 2021 and 2022.
+
+## Demograpic of Victims

### 2024-01-25 outputs/paper/paper.qmd
Final Draft

@@
 format: pdf
+fig-pos: "H"
 number-sections: true
@@
 
-Through R programming language [@R] we analyze the data obtained from Open data Toronto [Open data Toronto].
+Through R programming language [@R] we analyze the data obtained from Open data Toronto Portal [@rOpenDataToronto].
 This dataset provides a detailed account of all recorded count of victim under the jurisdiction of the Toronto Police Services over an eight-period.
 Our research offers an examination of cleaned data points and produced visual formats that reveal the shapes the crime type as well as their victims across different genders and age cohorts.
-The analysis, supplemented with various analytical tools like Tidyverse [@tidyverse], Janitor [@janitor], and Knitr [@knitr], delves deep into the trends, revealing patterns that could have implications for policy formulation and implementation.
+The analysis, supplemented with various analytical tools like Tidyverse [@rTidyverse], AER [@rAER], Janitor [@rJanitor], and Knitr [@rKnitr], delves deep into the trends, revealing patterns that could have implications for policy formulation and implementation.
 
@@
 
-The data used in this paper was collected by the OpenDataToronto Library [@citeODT]. The specific dataset used in this research is the
+The data used in this paper was collected by the OpenDataToronto Library [@rOpenDataToronto]. The specific dataset used in this research is the
 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset] from the year 2014 to 2022.
-The dataset is categorized by crime subtype against a person, so basically four-types: Assault, Robbery, Sexual Violation, and Other offenses.
+The dataset is categorized by crime against a person, so subtype includes: Assault, Robbery, Sexual Violation, and Other offenses.
 The stratification of the dataset follows victim demographics such as age group, age cohort and gender.
@@
 ## Data Cleaning
+
+Data used in this paper was downloaded, cleaned and analyzed with the programming language R [@R].
+Also with support of additional packages in R: `tidyverse` [@rTidyverse], `janitor` [@rJanitor], `dplyr` [@rDplyr], `readr` [@rReadr].
+
 ```{r}
@@
 
-Data used in this paper was downloaded, cleaned and analyzed with the programming language R [@citeR].
-Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `dplyr` [@citeDplyr], `readr` [@citeReadr], `knitr` [@citeknitr].
-The Table 1 showcases the cleaned dataset from the raw data extraced from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].
+The Table 1 showcases the cleaned dataset from the raw data extracted from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].
 
-# Results & Analysis
+The 'category' & 'assault subtype' columns from the raw data were irrelavant so removed due to their redundancy and lack of relevance to the focus of this study.
+No other data was discarded without careful consideration of its impact on the overall report.
+
+## Alternative Datasets
+
+Alternative datasets from other municipal or federal crime databases could have been utilized, but the Toronto Police Service's dataset was chosen for its specificity to the region of interest and for its detailed breakdown of victim demographics, which is central to our analysis.
+
+# Results {#sec-3}
 
@@
        x = "Year",
-       y = "Total Victims of Crime Over Year") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
-
+       y = "Total Victims of Crime") +
+  theme_minimal()
 ```
@@
 
-## Demograpic of Victims
+## Demographics of Victim
+
+### By Gender
+@fig-victims-by-Gender displays the distribution of victims of crimes grouped by gender across different years, from 2014 to 2022.
+The bars represent the total count of victims for each gender category, with separate colors for females (F), males (M), and an Unknown category (U).
+We can infer for every year from 2014-22 Female victims have higher count compared to Male victims.
+The data also showcase significantly lower number of victims as the remain ‘Unknown’.
+The figure high variation in terms of victims of crime over year.
+There isn’t any clear pattern in the data.A notable decrease in victim count for all genders can be seen around 2020.
+This could be an area of further investigation, as external factors such as the COVID-19 pandemic may have influenced crime rates and reporting.
+
+```{r}
+#| label: fig-victims-by-Gender
+#| fig-cap: Total Victims of Crime over Year (2014-2022)
+#| echo: false
+#| warning: false
+#| message: false
+victim_gender_year <- cleaned_dataset %>%
+  group_by(report_year, sex) %>%
+  summarise(total_victim_count = sum(victim_count)) %>%
+  ungroup()
+
+ggplot(victim_gender_year, aes(x = report_year, y = total_victim_count, fill = sex)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  labs(title = "Victims of Crime by Gender Across Years",
+       x = "Year",
+       y = "Total Victim Count",
+       fill = "Gender") +
+  theme_minimal()
+```
+
+### By Age Group
+
+Moving on, Victims belong to different groups and from the @tbl-Avg-Victim-of-crime-by-age-group.
+As depicted in Table 3, the sample summary of statistics by age group, we observed that the average victim of crime for Adults and Youth are relatively high compared to Children and Unknown age groups throughout the the 8-year period More Specifically, Adults had an average victims over the years has been the highest that is approximately 259, and even the Youths have considerably high, about 266 victims per year.
+In contrast, Children has low average no.of victims i.e.
+about 134 victims per year, and the Unknown category had the lowest victims.
+
+```{r}
+#| label: tbl-Avg-Victim-of-crime-by-age-group
+#| tbl-cap: Average Victims by Age Group (2014-2022)
+#| echo: false
+
+cleaned_dataset |>
+  group_by(age_group) |>
+  summarize(
+    mean_Victim_count = mean(victim_count),
+  ) |>
+  kable(
+    col.names = c("Crime", "Average Victim Count"),
+    booktabs = TRUE
+    )
+
+```
+
+Followed by @tbl-Avg-Victim-of-crime-by-age-group we created a line graph which provides visual represenation of the age group and the trend it follows from 2014-2022.
+So from @fig-Avg-Victims-2014-2022 it is oberved that the Adult and Youth age groups exhibit a higher and more volatile trend line over the years, indicating fluctuations in the number of victims.
+The Adult & Youth category, in particular, shows a peak around 2018, followed by a decline and a subsequent rise by 2022.
+On the other hand, the Child group's line remains fairly consistent with a slight increase over the years, and the Unknown category's line is the flattest, indicating the least variation in victim counts over the years but has remain consistent.
+```{r}
+#| label: fig-Avg-Victims-2014-2022
+#| fig-cap: The Average No. of Victims over eight-year period
+#| echo: false
+#| warning: false
+#| message: false
+
+# Assuming cleaned_dataset is your data frame and it has the columns 'report_year', 'age_group', and 'victim_count'
+average_victims_per_year_age_group <- cleaned_dataset %>%
+  group_by(report_year, age_group) %>%
+  summarize(avg_victim_count = mean(victim_count)) %>%
+  ungroup()
+
+# To visualize this data, you can create a line plot for each age group
+ggplot(average_victims_per_year_age_group, aes(x = report_year, y = avg_victim_count, group = age_group, color = age_group)) +
+  geom_line() +  # Add lines for each age group
+  geom_point() +  # Add points
+  labs(title = "Average Number of Victims per Year by Age Group",
+       x = "Year",
+       y = "Average Number of Victims",
+       color = "Age Group") +
+  theme_minimal() +
+  theme(legend.position = "bottom")  # Position the legend at the bottom
+
+```
+
+### By Age Cohort
+
+```{r}
+#| label: fig-Total-Victims-by-Age 
+#| fig-cap: Overall Victims by Age-Cohort & Crime categories in eight-year Period
+#| echo: false
+#| warning: false
+#| message: false
+
+# Summarize the total count by age cohort and subtype
+crimes_age_subtype <- cleaned_dataset %>%
+  group_by(age_cohort, subtype) %>%
+  summarise(total_count = sum(victim_count)) # Ensure grouped data is ungrouped after summarization
+
+# Create a horizontal bar graph for Crimes by Age Cohort and Subtype
+ggplot(crimes_age_subtype, aes(x = total_count, y = reorder(age_cohort, total_count), fill = subtype)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Victim of Crime by Age Cohort and Crime Subtype",
+       x = "Total Count",
+       y = "Age Cohort") +
+  theme_minimal() +
+  theme(legend.position = "bottom") # Adjust legend position if necessary
+
+```
+@fig-Total-Victims-by-Age More than half of victims between the age cohort of '25 to 34' has the highest total count of crime victims across all crime categories, indicating this age group has the most affected  victims of a crime.
+The '18 to 24' and '35 to 44' cohorts also have significant total counts of victims, with assault being the predominant crime subtype.
+This may reflect a vulnerability or exposure to crime among these younger adult age groups.
+The '45 to 54' and '55 to 64' age cohorts show moderate victim counts, with assault still being the most common subtype, but with a noticeable presence of other crime types as well.
+The '65+' and '<12' age groups have lower total counts of crime victims, indicating a lower incidence of victimization in the oldest and youngest cohorts.
+However, it's important to note that the impact of crime on these groups can be particularly profound.
+The '12 to 17' cohort, representing youth, shows a considerable count of victims, predominantly in the assault subtype.
+This highlights the importance of focusing on crime prevention strategies for adolescents.
+The 'Unknown' category has the least total count, indicating that these may be incidents where the victim's age was not reported or could not be reported or determine
+
+# Discussion {#sec-4}
+
+Our comprehensive analysis of crime data from 2014 to 2022 in Toronto has revealed distinct patterns of victimization across various demographics.
+The analysis conducted in the Result section highlights a worrying trend of increased victimization up to 2019 which implies a wave of increase in crime the criminal activity in the city of Toronto.
+The followed by a sudden decline the victims number in 2020, and a resurgence thereafter.
+This pattern could be influenced by the global pandemic, underscoring the interplay between societal factors and crime rates.
+The data align with broader trends noted in the literature, including the risk among young adults.
+Also The victims by gender showcase female have higher threats as majority of time the victims tend to be women.
+The comparative victimization rates between genders reveal entrenched patterns that necessitate a gender-responsive approach to crime prevention.
+As the @fig-victims-by-Gender, and age, detailed in @tbl-Avg-Victim-of-crime-by-age-group and @fig-Avg-Victims-2014-2022.
+This requires a prompt reevaluation in the strategies of targeted crime prevention policies.
+The victim of the '25 to 34' age cohort demands a policy response as well as research under the high victim count can help address the vulnerabilities of this demographic.
+
+While our report provides a molecular insight into crime’s victim.
+However, there exist some limitations, such as the 'Unknown' category signifies gaps in data collection.
+Moreover, the categorization of crimes and the category ‘Other’ are areas where data refinement could spot some light on.
+The use of police-reported data does not showcase the motive or the relationship between victim and criminal, as not all incidents are reported.
+The insight does not represent the complexity of individual cases.
+
+# Conclusion {#sec-5}
+
+In conclusion to our report, Victims of Crime in Toronto the years of 2014 to 2022 has highlighted evident patterns and trends, most particularly the disproportionate impact of crime on the victims of age 25 to 34-year-old.
+The analysis indicates that young adults face the highest rates of victimization, while there is a notable decrease in reported victims during the pandemic year of 2020, hinting at the influence of external factors on crime trends.
+Gender disparities in victimization were also evident, with females consistently representing a higher percentage of victims.
+However there is also need for improved data collection and explore new approach to crime prevention that will address vulnerabilities & public safety policies of different demographic segments.
+
+# Referneces {#sec-6}

### 2024-01-25 outputs/paper/paper.qmd
Final Draft

@@
 date-format: long
-abstract: "Public Safety and security are a paramount concern in the 20th century, the analysis of statistics in Victimology helps to visualize and derive the conclusions of urban crime dynamics.
-This analytical report delves into the trends of victims of a certain crime in Toronto from the year 2014 to 2022.
-Through a series of data visualization approach to examine the relationship between crime incidents and demographic factors.
+abstract: "Public Safety and security are a paramount concern in the 21st century, the analysis of statistics in Victimology helps to visualize and derive the conclusions of urban crime dynamics.
+This analytical report delves into the trends of victims of a crime in Toronto from the year 2014 to 2022.
+This report employs a series of data visualization approaches to examine the relationship between crime incidents and demographic factors.
 The data reveals significant fluctuations in crime rates in correspond to the socio-economic shifts, including a notable dip in crimes reported during pandemic between 2019-2020.
@@
 fig-pos: "H"
+toc: true
 number-sections: true
@@
 # Introduction
-The concern of Public safety in the Urban society is a necessary step towards safeguarding people of that community.
-The city of Toronto has various socio-economic challenges one of them being Crime against a person.
+
+Concern of public safety in the urban societies is a necessary step towards safeguarding the residents of that community.
+The city of Toronto has various socio-economic challenges one of them being crime against a person.
 A crime incident might leave a victim and our report aims to explore the patterns and trend of crimes against the person.
-The paper aims to understand how this trends have evolved among victims and analyzing there  demograpics from 2014 to 2022.
-To elaborate, the report context is not only quantifying crime victims but also interpret and explain whereby it motivates to a better approach towards urban safety and policies.
-The Criminal victimization is a frightening and unsettling experience for many Canadians (Cite).
-
+The paper aims to understand how this trends have evolved among victims and to analyze their demographics from 2014 to 2022.
+Furthermore, it's important to note that the report's scope extends beyond merely quantifying crime victims but it also interpret and explain whereby it motivates to a better approach towards urban safety and policies.
+The Criminal victimization is a frightening and unsettling experience for many Canadians (@statscanada).
 Victimology is a vast sub-category of criminology to understand various emotional factors and its causation, however based on our police annual report – victims of  crime showcase a statistical record based on which the report will share the knowledge to work on the gap to understand between victim demographics and crime.
-This report addresses this gap by visualizing crime statistics through a demographic lens, thereby unveiling patterns that may not be evident in numerical data alone
-
-Through R programming language [@R] we analyze the data obtained from Open data Toronto Portal [@rOpenDataToronto].
+This report addresses this gap by visualizing crime statistics through a demographic lens, thereby unveiling patterns that may not be evident in numerical data alone.
+Through R programming language [@R], in this report, we analyze the data obtained.
+from Open data Toronto Portal [@rOpenDataToronto].
 This dataset provides a detailed account of all recorded count of victim under the jurisdiction of the Toronto Police Services over an eight-period.
@@
 The analysis, supplemented with various analytical tools like Tidyverse [@rTidyverse], AER [@rAER], Janitor [@rJanitor], and Knitr [@rKnitr], delves deep into the trends, revealing patterns that could have implications for policy formulation and implementation.
-
-The Introduction is followed by the Data section.
+The introduction is followed by the data section.
 Within the data section we talk about the data source used for this report, the data cleaning process applied and the result that is the visual analysis of trends and pattern involved.
 The Discussion section shares additional insights into the data findings, as well as discussing measurement errors and limitations of this analysis.
-Lastly, the document summarizes the analysis.
+Lastly, the document summarizes the analysis with conclusion.
 
@@
 
-The data used in this paper was collected by the OpenDataToronto Library [@rOpenDataToronto]. The specific dataset used in this research is the
+The data used in this paper were collected by the OpenDataToronto Library [@rOpenDataToronto]. The specific dataset used in this research is the
 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset] from the year 2014 to 2022.
-The dataset is categorized by crime against a person, so subtype includes: Assault, Robbery, Sexual Violation, and Other offenses.
-The stratification of the dataset follows victim demographics such as age group, age cohort and gender.
+The dataset is categorized by crime against a person, so subtype includes: assault, robbery, sexual Violation, and other offenses.
+The stratification of the dataset follows victim demographics such as age group, age cohort, and gender.
 
@@
 
-The Table 1 showcases the cleaned dataset from the raw data extracted from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].
+The Table 1 (@tbl-Cleaned-data-output) showcases the cleaned dataset from the raw data extracted from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].
 
@@
 
-# Use kable from knitr to create a simple table
+# Making table from Kable for pdf paper
 
@@
 
-# Creating the line graph
+# Creating the line graph showing of the Year & Total Victim over 8 year period
 ggplot(summarized_data, aes(x = report_year, y = Total_Count)) +
@@
 There isn’t any clear pattern in the data.A notable decrease in victim count for all genders can be seen around 2020.
-This could be an area of further investigation, as external factors such as the COVID-19 pandemic may have influenced crime rates and reporting.
+This could be an area of further investigation, as external factors such as the COVID-19 pandemic may have influenced crime rates and reports conducted during that year.
 
@@
 #| echo: false
-
+# Using Cleaned Data set and grouping vaiable of interest
 cleaned_dataset |>
@@
 
-# Assuming cleaned_dataset is your data frame and it has the columns 'report_year', 'age_group', and 'victim_count'
+# Using cleaned_dataset that has the columns 'report_year', 'age_group', and 'victim_count'
 average_victims_per_year_age_group <- cleaned_dataset %>%
@@
 
-# To visualize this data, you can create a line plot for each age group
+# To visualize - a line graph for each age group
 ggplot(average_victims_per_year_age_group, aes(x = report_year, y = avg_victim_count, group = age_group, color = age_group)) +
-  geom_line() +  # Add lines for each age group
-  geom_point() +  # Add points
+  geom_line() +  
+  geom_point() +  
   labs(title = "Average Number of Victims per Year by Age Group",
@@
   theme_minimal() +
-  theme(legend.position = "bottom")  # Position the legend at the bottom
+  theme(legend.position = "bottom")
 
@@
 
-# Summarize the total count by age cohort and subtype
+# Making subtables for graph
 crimes_age_subtype <- cleaned_dataset %>%
   group_by(age_cohort, subtype) %>%
-  summarise(total_count = sum(victim_count)) # Ensure grouped data is ungrouped after summarization
+  summarise(total_count = sum(victim_count))
 
-# Create a horizontal bar graph for Crimes by Age Cohort and Subtype
+# Horizontal bar graph for Crimes and its victim Age Cohort and Subtype
 ggplot(crimes_age_subtype, aes(x = total_count, y = reorder(age_cohort, total_count), fill = subtype)) +
@@
   theme_minimal() +
-  theme(legend.position = "bottom") # Adjust legend position if necessary
+  theme(legend.position = "bottom")
 

### 2024-01-25 outputs/paper/paper.qmd
Final Draft

@@
 The analysis, supplemented with various analytical tools like Tidyverse [@rTidyverse], AER [@rAER], Janitor [@rJanitor], and Knitr [@rKnitr], delves deep into the trends, revealing patterns that could have implications for policy formulation and implementation.
-The introduction is followed by the data section.
-Within the data section we talk about the data source used for this report, the data cleaning process applied and the result that is the visual analysis of trends and pattern involved.
-The Discussion section shares additional insights into the data findings, as well as discussing measurement errors and limitations of this analysis.
-Lastly, the document summarizes the analysis with conclusion.
+Following this introduction, the Data section discusses our data sources, cleaning process, and the resultant visual analysis.
+The Discussion section delves into the implications of our findings, addressing potential measurement errors and limitations.
+Finally, the paper concludes with a summary of our analysis and its implications for policy formulation
 

### 2024-01-25 outputs/paper/paper.qmd
Final Commit

@@
 date-format: long
-abstract: "Public Safety and security are a paramount concern in the 21st century, the analysis of statistics in Victimology helps to visualize and derive the conclusions of urban crime dynamics.
+abstract: "Public safety and security are a paramount concern in the 21st century, the analysis of statistics in victimology helps to visualize and derive the conclusions of urban crime dynamics.
 This analytical report delves into the trends of victims of a crime in Toronto from the year 2014 to 2022.
@@
 Furthermore, it's important to note that the report's scope extends beyond merely quantifying crime victims but it also interpret and explain whereby it motivates to a better approach towards urban safety and policies.
-The Criminal victimization is a frightening and unsettling experience for many Canadians (@statscanada).
+The criminal victimization is a frightening and unsettling experience for many Canadians (@statscanada).
 Victimology is a vast sub-category of criminology to understand various emotional factors and its causation, however based on our police annual report – victims of  crime showcase a statistical record based on which the report will share the knowledge to work on the gap to understand between victim demographics and crime.
@@
 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset] from the year 2014 to 2022.
-The dataset is categorized by crime against a person, so subtype includes: assault, robbery, sexual Violation, and other offenses.
+The dataset is categorized by crime against a person, so subtype includes: assault, robbery, sexual violation, and other offenses.
 The stratification of the dataset follows victim demographics such as age group, age cohort, and gender.
@@
 
-The Table 1 (@tbl-Cleaned-data-output) showcases the cleaned dataset from the raw data extracted from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].
+The @tbl-Cleaned-data-output showcases the cleaned dataset from the raw data extracted from the 'Police-Annual-Statistical-Report-Victims-of-Crimes' [@citeDataset].
 
@@
 
-Alternative datasets from other municipal or federal crime databases could have been utilized, but the Toronto Police Service's dataset was chosen for its specificity to the region of interest and for its detailed breakdown of victim demographics, which is central to our analysis.
+Alternative datasets from other municipal or federal crime databases for countries like Canada or US which could have been utilized, but the Toronto Police Service's dataset was chosen for its specificity to the region of interest and for its detailed breakdown of victim demographics, which is central to our analysis.
 
@@
 This may reflect a vulnerability or exposure to crime among these younger adult age groups.
-The '45 to 54' and '55 to 64' age cohorts show moderate victim counts, with assault still being the most common subtype, but with a noticeable presence of other crime types as well.
+The '45 to 54' and '55 to 64' age cohorts show moderate victim counts, with assault still being the most common crime , but with a noticeable presence of other crime types as well.
 The '65+' and '<12' age groups have lower total counts of crime victims, indicating a lower incidence of victimization in the oldest and youngest cohorts.
 However, it's important to note that the impact of crime on these groups can be particularly profound.
-The '12 to 17' cohort, representing youth, shows a considerable count of victims, predominantly in the assault subtype.
-This highlights the importance of focusing on crime prevention strategies for adolescents.
-The 'Unknown' category has the least total count, indicating that these may be incidents where the victim's age was not reported or could not be reported or determine
+The '12 to 17' cohort, representing youth, shows a considerable count of victims, predominantly in the assault crime.
+This highlights the importance of focusing on crime prevention strategies for teenage students.
+The 'Unknown' category has the least total count, indicating that these may be incidents where the victim's age was not reported.
 

### 2024-01-25 outputs/paper/paper.qmd
Final Commit

@@
 author: "Mohammed Yusuf Shaikh"
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/Yusuf365/project_file_data_analyisis_1.git"
 date: today
